---
title: "What does the Data Team do?"
subtitle: "<br/>Featuring R"
author: "Patrick Anker"
institute: "Global TIES for Children"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---
class: inverse, middle, center

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(xaringanthemer)
style_mono_accent(
  text_font_google = google_font("Montserrat", "300", "300i"),
  code_font_google = google_font("Fira Mono")
)

library(magrittr)
```

# Before we begin...

All materials in this presentation can be downloaded from 

## Data Team Motivation

- Studies feature multiple kinds of data (surveys, activity trackers, etc.)
- Varying degree of data trustworthiness and cleanliness
- Previously no 3EA-specific standards for data management

---

## Why R?

- Needed flexible language to address data problems
- Easy to use

--

- Mayarí started the team, and she knew R so... here we are lol

---
class: inverse, middle, center

# Basic R Setup

---

## Installation

- [Install R](https://cran.r-project.org/)
- [Install RStudio](https://rstudio.com/products/rstudio/download/)

We need to install both, unlike STATA for example, since R does not provide an _integrated development environment_ (IDE). That's where RStudio comes in.

<img src="images/r-logo.svg" width="20%" style="float:left" /> <img src="images/rstudio-logo.svg" width="40%" style="float:right" />

---

## Packages

Packages collections of code that extend base R's functionality.

```{r, echo=FALSE}
set.seed(1572)

installed.packages() %>% 
  tibble::as_tibble() %>% 
  dplyr::select(Package, Version) %>% 
  dplyr::slice_sample(n = 9)
```

---

## Packages

Let's install some popular and handy packages. Open RStudio, and run these lines in the "Console" tab: 

```{r, eval=FALSE}
install.packages("tidyverse")
install.packages("tidymodels")
```

You only need to run these **once**. Once you have them installed, you don't need to keep reinstalling them (I see this often in R analysis scripts!!).
